{\bf Practical Concurrency Testing} \\ \normalsize \vspace{1em} \begin{tabular}{c} {\em or: How I Learned to Stop Worrying and Love the Exponential Explosion} \end{tabular}

### 2018-12-23 thesis/intro.tex
やはり data race analysis

\begin{itemize}
	\item {\bf Background:} \Cref{chap:background} will present the requisite background material on concurrent programming, stateless model checking, and the various types of programs targeted by Landslide,
		\revisionminor{concluding with a glossary of terminology for the reader's convenient reference.}
		%(\cref{sec:glossary})
	\item {\bf Landslide:} \Cref{chap:landslide} explains the design and implementation of Landslide
		%the core of my stateless model checking framework,
		and all the special features it's been equipped with over the years.
		It is the foundation upon which all three contributions above build.
	\item {\bf Quicksand:} \Cref{chap:quicksand} introduces
		\revisionminor{Quicksand,
		a wrapper framework for Landslide
		which incorporates data-race analysis at run-time to find new preemption points
		and uses the new Iterative Deepening algorithm to}
		more intelligently choose which state spaces to test, corresponding to contribution 1 above.
	\item {\bf Education:} \Cref{chap:education} discusses my evaluation of Landslide
		in CMU's
		\revisionminor{and The Pennsylvania State University's
		class environment using thread libraries based on the Pebbles kernel},
		and in the University of Chicago's
		OS class environment using the Pintos kernel,
		corresponding to contribution 2 above.
	\item {\bf Transactions:} \Cref{chap:tm} presents my extension of Landslide's concurrency model to handle transactional concurrency and the evaluation thereof, corresponding to contribution 3 above.
	\item {\bf Related Work:} \Cref{chap:relatedwork} honors my neighbours and ancestors in research spirit.
	\item {\bf Future Work:} \Cref{chap:warpzone} identifies stateless model checking's remaining shortcomings and how new research might address them.
	\item {\bf Conclusion:} \Cref{chap:conclusion} offers some thoughts on the future of the field.
\end{itemize}

>>>>

\begin{itemize}
	\item {\bf Background:} \Cref{chap:background} will present the requisite background material on concurrent programming, stateless model checking, and the various types of programs targeted by Landslide,
		\revisionminor{concluding with a glossary of terminology for the reader's convenient reference.}
		%(\cref{sec:glossary})
	\item {\bf Landslide:} \Cref{chap:landslide} explains the design and implementation of Landslide
		%the core of my stateless model checking framework,
		and all the special features it's been equipped with over the years.
		It is the foundation upon which all three contributions above build.
	\item {\bf Quicksand:} \Cref{chap:quicksand} introduces
		\revisionminor{Quicksand,
		a wrapper framework for Landslide
		which incorporates data race analysis at run-time to find new preemption points
		and uses the new Iterative Deepening algorithm to}
		more intelligently choose which state spaces to test, corresponding to contribution 1 above.
	\item {\bf Education:} \Cref{chap:education} discusses my evaluation of Landslide
		in CMU's
		\revisionminor{and The Pennsylvania State University's
		class environment using thread libraries based on the Pebbles kernel},
		and in the University of Chicago's
		OS class environment using the Pintos kernel,
		corresponding to contribution 2 above.
	\item {\bf Transactions:} \Cref{chap:tm} presents my extension of Landslide's concurrency model to handle transactional concurrency and the evaluation thereof, corresponding to contribution 3 above.
	\item {\bf Related Work:} \Cref{chap:relatedwork} honors my neighbours and ancestors in research spirit.
	\item {\bf Future Work:} \Cref{chap:warpzone} identifies stateless model checking's remaining shortcomings and how new research might address them.
	\item {\bf Conclusion:} \Cref{chap:conclusion} offers some thoughts on the future of the field.
\end{itemize}

### 2018-12-23 thesis/landslide.tex
Data-race (haha)
やはり data race analysis

\begin{itemize}
	\item {\tt -C}: enable ``control experiment'' mode, i.e., run only 1 instance of Landslide, with all (non-data-race) preemption points enabled in advance
	\item {\tt -I}: enable Iterative Context Bounding (requires {\tt -C}, although future work may relax this restriction);
		this generally causes bugs to be found faster should they exist, but degrades completion time
		(\cref{sec:landslide-icb})
	\item {\tt -0}: enable Preempt-Everywhere mode
		(\cref{sec:quicksand-impl-modes}, \cref{sec:quicksand-impl-landslide}, requires {\tt -C})
	\item {\tt -M}: enable Maximal State Space mode,
		which prioritizes the maximal state space to optimize for fast verification,
		abandoning all subset jobs even if they might find bugs faster
		(\cref{sec:tm-eval}, incompatible with {\tt -C}).
		According to \cref{sec:quicksand-soundness}'s soundness proofs,
		this is equivalent to {\tt -0}
		(and according to my experience, \revisionminor{much} faster as well).
	\item {\tt -H}: use Limited Happens-Before for data-race analysis (\cref{sec:background-hb})
		(default for Pebbles kernelspace mode)
	\item {\tt -V}: use vector-clock-based Pure Happens-Before for data-race analysis (\cref{sec:background-hb})
		(default for P2/PSU userspace and Pintos modes)
	\item {\tt -X}: support transactional memory (\Cref{chap:tm})
	\item {\tt -A}: support multiple abort codes during transaction failure (\cref{sec:tm-implementation});
		required for testing programs which behave differently under different abort circumstances,
		but impacts the state space size
	\item {\tt -S}: suppress retry aborts during transaction failure (\cref{sec:tm-implementation})
	\item {\tt -R}: enable retry-set state space reduction for transactional tests (\cref{sec:tm-implementation})
	\item {\tt -P}: support Pintos architecture (enabled automatically when {\tt pintos-setup.sh} is run)
	\item {\tt -4}: support Pebbles architecture (enabled automatically when either {\tt p2-setup.sh} or {\tt psu-setup.sh} is run)
	\item {\tt -e ETAFACTOR}: configure heuristic state space ETA deferring factor (\cref{sec:quicksand-id}; described in detail in {\tt id/option.c})
	\item {\tt -E ETATHRESH}: configure heuristic threshold of state space progress for judging ETA stability (\cref{sec:quicksand-id}; described in detail in {\tt id/option.c})
\end{itemize}

@@

\begin{itemize}
	\item Extends Preemption Sealing for finer-grained control
		over whether individual functions and/or global variables
		should be considered for, or excluded from,
		DPOR and/or data-race analysis
		(\cref{sec:landslide-config-landslide};
		{\tt thrlib\_function} and {\tt ignore\_dr\_function}, respectively).
		This allows a test case author to precisely target the exact subset of code
		she is interested in testing,
		avoiding unnecessary state space inflation from trusted code.
	\item Adds a mutex testing mode to configure whether the
		data race analysis should consider locks' internal accesses
		(\cref{sec:landslide-dynamicconfig}; {\tt TESTING\_MUTEXES}).
		This allows the user to verify that her lock implementation
		correctly provides mutual exclusion
		(upcoming in \cref{sec:education-pebbles-tests}; {\tt mutex\_test}),
		before relying on them as trusted code in subsequent tests.
	\item Implements the sleep set extension to DPOR \cite{partial-order-methods,dpor,optimal-dpor},
		which identifies and skips further equivalences based on which memory accesses
		DPOR intended to reorder,
		%, rather than just switching threads and seeing what happens
		in a way that is slightly less optimal compared to prior work,
		but much simpler to implement
		and requires less state at runtime (\cref{sec:landslide-sleepsets}).
	\item Heuristically identifies infinite loops and livelocks
		by leveraging knowledge of how long the test program should be expected to run
		obtained during previous executions,
		uniquely suited to stateless model checking (\cref{sec:landslide-infloop}).
	\item Provides automatic code instrumentation facilities,
		beyond just the simple task of identifying important function boundaries,
		that attempt to understand a wide variety of even the most unusual student implementations
		(\cref{sec:landslide-pebblesspecifics},
		\cref{sec:landslide-pintosspecifics}).
		This includes an advanced heuristic for detecting ad-hoc synchronization loops
		that Landslide should treat as blocking, %a thread,
		even if it does not trigger the automatically-inserted blocking annotations
		(\cref{sec:landslide-blocking-yield}),
		and a further heuristic to ensure that any inaccuracy on the part of the first heuristic
		does not result in false-positive bug reports (\cref{sec:landslide-fp-deadlock}).
\end{itemize}

>>>>

\begin{itemize}
	\item {\tt -C}: enable ``control experiment'' mode, i.e., run only 1 instance of Landslide, with all (non-data-race) preemption points enabled in advance
	\item {\tt -I}: enable Iterative Context Bounding (requires {\tt -C}, although future work may relax this restriction);
		this generally causes bugs to be found faster should they exist, but degrades completion time
		(\cref{sec:landslide-icb})
	\item {\tt -0}: enable Preempt-Everywhere mode
		(\cref{sec:quicksand-impl-modes}, \cref{sec:quicksand-impl-landslide}, requires {\tt -C})
	\item {\tt -M}: enable Maximal State Space mode,
		which prioritizes the maximal state space to optimize for fast verification,
		abandoning all subset jobs even if they might find bugs faster
		(\cref{sec:tm-eval}, incompatible with {\tt -C}).
		According to \cref{sec:quicksand-soundness}'s soundness proofs,
		this is equivalent to {\tt -0}
		(and according to my experience, \revisionminor{much} faster as well).
	\item {\tt -H}: use Limited Happens-Before for data race analysis (\cref{sec:background-hb})
		(default for Pebbles kernelspace mode)
	\item {\tt -V}: use vector-clock-based Pure Happens-Before for data race analysis (\cref{sec:background-hb})
		(default for P2/PSU userspace and Pintos modes)
	\item {\tt -X}: support transactional memory (\Cref{chap:tm})
	\item {\tt -A}: support multiple abort codes during transaction failure (\cref{sec:tm-implementation});
		required for testing programs which behave differently under different abort circumstances,
		but impacts the state space size
	\item {\tt -S}: suppress retry aborts during transaction failure (\cref{sec:tm-implementation})
	\item {\tt -R}: enable retry-set state space reduction for transactional tests (\cref{sec:tm-implementation})
	\item {\tt -P}: support Pintos architecture (enabled automatically when {\tt pintos-setup.sh} is run)
	\item {\tt -4}: support Pebbles architecture (enabled automatically when either {\tt p2-setup.sh} or {\tt psu-setup.sh} is run)
	\item {\tt -e ETAFACTOR}: configure heuristic state space ETA deferring factor (\cref{sec:quicksand-id}; described in detail in {\tt id/option.c})
	\item {\tt -E ETATHRESH}: configure heuristic threshold of state space progress for judging ETA stability (\cref{sec:quicksand-id}; described in detail in {\tt id/option.c})
\end{itemize}

@@

\begin{itemize}
	\item Extends Preemption Sealing for finer-grained control
		over whether individual functions and/or global variables
		should be considered for, or excluded from,
		DPOR and/or data race analysis
		(\cref{sec:landslide-config-landslide};
		{\tt thrlib\_function} and {\tt ignore\_dr\_function}, respectively).
		This allows a test case author to precisely target the exact subset of code
		she is interested in testing,
		avoiding unnecessary state space inflation from trusted code.
	\item Adds a mutex testing mode to configure whether the
		data race analysis should consider locks' internal accesses
		(\cref{sec:landslide-dynamicconfig}; {\tt TESTING\_MUTEXES}).
		This allows the user to verify that her lock implementation
		correctly provides mutual exclusion
		(upcoming in \cref{sec:education-pebbles-tests}; {\tt mutex\_test}),
		before relying on them as trusted code in subsequent tests.
	\item Implements the sleep set extension to DPOR \cite{partial-order-methods,dpor,optimal-dpor},
		which identifies and skips further equivalences based on which memory accesses
		DPOR intended to reorder,
		%, rather than just switching threads and seeing what happens
		in a way that is slightly less optimal compared to prior work,
		but much simpler to implement
		and requires less state at runtime (\cref{sec:landslide-sleepsets}).
	\item Heuristically identifies infinite loops and livelocks
		by leveraging knowledge of how long the test program should be expected to run
		obtained during previous executions,
		uniquely suited to stateless model checking (\cref{sec:landslide-infloop}).
	\item Provides automatic code instrumentation facilities,
		beyond just the simple task of identifying important function boundaries,
		that attempt to understand a wide variety of even the most unusual student implementations
		(\cref{sec:landslide-pebblesspecifics},
		\cref{sec:landslide-pintosspecifics}).
		This includes an advanced heuristic for detecting ad-hoc synchronization loops
		that Landslide should treat as blocking, %a thread,
		even if it does not trigger the automatically-inserted blocking annotations
		(\cref{sec:landslide-blocking-yield}),
		and a further heuristic to ensure that any inaccuracy on the part of the first heuristic
		does not result in false-positive bug reports (\cref{sec:landslide-fp-deadlock}).
\end{itemize}

### 2018-12-23 thesis/overview.tex
Data-race (haha)
やはり data race analysis

\revisionminor{This chapter introduces the necessary background material on concurrency
(\cref{sec:overview-concurrency}),
stateless model checking,
(\cref{sec:overview-stateless}),
data-race analysis
(\cref{sec:background-datarace}),
and the relevant undergraduate operating systems classes
(\cref{sec:overview-edu}).
It concludes with a glossary of terminology
(\cref{sec:glossary})
that I hope the reader may find convenient as a quick reference
while reading later chapters as well.}

@@

Each such abstraction provides certain semantics about \revisionminor{which} thread interleavings can arise surrounding their use.
When building a tool for testing concurrent programs,
one may include some computational understanding of the behaviour of any, or all, of these abstractions.
Annotating a certain abstraction's semantics treats it as a trusted concurrency primitive in its own right,
and allows the testing tool to reduce the possible space of interleavings
(or the set of false positive data-race candidates reported, et cetera;
\revisionminor{see \cref{sec:landslide-config-landslide}})
at the cost of increasing the implementation and theoretical complexity of the analysis
\revisionminor{(\cref{sec:tm-abstraction})}.
This thesis will consider locks, descheduling, and transaction begin/end as the only concurrency primitives,
and assume the others listed above are implemented using those as building blocks (an exercise for the reader \cite{thrlib}).

@@

\begin{figure}[h]
        \small
	\begin{center}
\begin{tabular}{c}
\begin{tabular}{rll}
	\\
	& \multicolumn{2}{c}{\texttt{\ctype{int} x = \const{0}; \ctype{bool} y = \const{false}; \ctype{mutex\_t} mx;}} \\
        \\
        & {\bf Thread 1} & {\bf Thread 2} \\
	1 & \texttt{\hilight{brickred}{x++;}~\ccomment{// A1}} & \\
	2 & \texttt{\call{mutex\_lock}(\&mx);} & \\
	3 & \texttt{\call{mutex\_unlock}(\&mx);} & \\
	4 & & \texttt{\call{mutex\_lock}(\&mx);} \\
	5 & & \texttt{\call{mutex\_unlock}(\&mx);} \\
	6 & & \texttt{\hilight{brickred}{x++;}~\ccomment{// A2}} \\
\end{tabular}
\\
\\
	{\normalsize (a) True potential data race.}
\\
\\
\begin{tabular}{rll}
        %& \multicolumn{2}{c}{\texttt{int x = 0; bool y = false; mutex\_t mx;}} \\
        & {\bf Thread 1} & {\bf Thread 2} \\
	1 & \texttt{\hilight{brickred}{x++;}~\ccomment{// B1}} & \\
	2 & \texttt{\call{mutex\_lock}(\&mx);} & \\
	3 & \texttt{y = \const{true};} & \\
	4 & \texttt{\call{mutex\_unlock}(\&mx);} & \\
	5 & & \texttt{\call{mutex\_lock}(\&mx);} \\
	6 & & \texttt{\ctype{bool} tmp = y;} \\
	7 & & \texttt{\call{mutex\_unlock}(\&mx);} \\
	8 & & \texttt{\flow{if} (tmp) \hilight{brickred}{x++;}~\ccomment{// B2}} \\
        %8 & & \texttt{if (tmp)} \\
        %9 & & \texttt{~~~~\hilight{brickred}{x++;}~// B2} \\
\end{tabular}
\\
\\
{\normalsize (b) No data race in any interleaving.}
\end{tabular}
	\end{center}
	\caption[Data-race analyses may be prone to either {\em false negatives} or {\em false positives}.]
		{{Data-race analyses may be prone to either {\em false negatives} or {\em false positives}.
	Applying Happens-Before to program (a) will miss the potential race possible between A1/A2 in an alternate interleaving,
	while using Limited Happens-Before on (b) will produce a false alarm on B1/B2.}}
\label{fig:hb-example}
\end{figure}

A data-race analysis may be either {\em static} (inspecting source code) \cite{racerx}
or {\em dynamic} (tracking individual accesses arising at run-time) \cite{tsan}.
This paper focuses exclusively on dynamic analysis,
so although our example refers to numbered source lines for ease of explanation,
in practice we are actually classifying the individual memory access events corresponding to those lines during execution.
Actually, each {\tt x++} statement likely compiles to two separate load or store instructions, so each of those two instructions from each of the two marked source lines pairwise will race (except for the two loads, which are both reads).

@@

Landslide implements both Happens-Before (henceforth referred to as {\em Pure Happens-Before} for clarity) and Limited Happens-Before.
\Cref{chap:quicksand} includes a comparison of the two approaches for the purpose of finding new preemption points for model checking.
%, we use the Limited Happens-Before relation for our analysis.
%The justification for this is that, while stand-alone data-race analyses must avoid inundating the user with false alarms \cite{racerx},
%my work incorporates data-race analysis in an internal feedback loop, and reports only directly observed failures to the user.
%Hence, I accept some overhead from false positives for the sake of more thorough testing.

@@

		that Landslide be only a data-race detector,

>>>>

\revisionminor{This chapter introduces the necessary background material on concurrency
(\cref{sec:overview-concurrency}),
stateless model checking,
(\cref{sec:overview-stateless}),
data race analysis
(\cref{sec:background-datarace}),
and the relevant undergraduate operating systems classes
(\cref{sec:overview-edu}).
It concludes with a glossary of terminology
(\cref{sec:glossary})
that I hope the reader may find convenient as a quick reference
while reading later chapters as well.}

@@

Each such abstraction provides certain semantics about \revisionminor{which} thread interleavings can arise surrounding their use.
When building a tool for testing concurrent programs,
one may include some computational understanding of the behaviour of any, or all, of these abstractions.
Annotating a certain abstraction's semantics treats it as a trusted concurrency primitive in its own right,
and allows the testing tool to reduce the possible space of interleavings
(or the set of false positive data race candidates reported, et cetera;
\revisionminor{see \cref{sec:landslide-config-landslide}})
at the cost of increasing the implementation and theoretical complexity of the analysis
\revisionminor{(\cref{sec:tm-abstraction})}.
This thesis will consider locks, descheduling, and transaction begin/end as the only concurrency primitives,
and assume the others listed above are implemented using those as building blocks (an exercise for the reader \cite{thrlib}).

@@

\begin{figure}[h]
        \small
	\begin{center}
\begin{tabular}{c}
\begin{tabular}{rll}
	\\
	& \multicolumn{2}{c}{\texttt{\ctype{int} x = \const{0}; \ctype{bool} y = \const{false}; \ctype{mutex\_t} mx;}} \\
        \\
        & {\bf Thread 1} & {\bf Thread 2} \\
	1 & \texttt{\hilight{brickred}{x++;}~\ccomment{// A1}} & \\
	2 & \texttt{\call{mutex\_lock}(\&mx);} & \\
	3 & \texttt{\call{mutex\_unlock}(\&mx);} & \\
	4 & & \texttt{\call{mutex\_lock}(\&mx);} \\
	5 & & \texttt{\call{mutex\_unlock}(\&mx);} \\
	6 & & \texttt{\hilight{brickred}{x++;}~\ccomment{// A2}} \\
\end{tabular}
\\
\\
	{\normalsize (a) True potential data race.}
\\
\\
\begin{tabular}{rll}
        %& \multicolumn{2}{c}{\texttt{int x = 0; bool y = false; mutex\_t mx;}} \\
        & {\bf Thread 1} & {\bf Thread 2} \\
	1 & \texttt{\hilight{brickred}{x++;}~\ccomment{// B1}} & \\
	2 & \texttt{\call{mutex\_lock}(\&mx);} & \\
	3 & \texttt{y = \const{true};} & \\
	4 & \texttt{\call{mutex\_unlock}(\&mx);} & \\
	5 & & \texttt{\call{mutex\_lock}(\&mx);} \\
	6 & & \texttt{\ctype{bool} tmp = y;} \\
	7 & & \texttt{\call{mutex\_unlock}(\&mx);} \\
	8 & & \texttt{\flow{if} (tmp) \hilight{brickred}{x++;}~\ccomment{// B2}} \\
        %8 & & \texttt{if (tmp)} \\
        %9 & & \texttt{~~~~\hilight{brickred}{x++;}~// B2} \\
\end{tabular}
\\
\\
{\normalsize (b) No data race in any interleaving.}
\end{tabular}
	\end{center}
	\caption[Data race analyses may be prone to either {\em false negatives} or {\em false positives}.]
		{{Data race analyses may be prone to either {\em false negatives} or {\em false positives}.
	Applying Happens-Before to program (a) will miss the potential race possible between A1/A2 in an alternate interleaving,
	while using Limited Happens-Before on (b) will produce a false alarm on B1/B2.}}
\label{fig:hb-example}
\end{figure}

A data race analysis may be either {\em static} (inspecting source code) \cite{racerx}
or {\em dynamic} (tracking individual accesses arising at run-time) \cite{tsan}.
This paper focuses exclusively on dynamic analysis,
so although our example refers to numbered source lines for ease of explanation,
in practice we are actually classifying the individual memory access events corresponding to those lines during execution.
Actually, each {\tt x++} statement likely compiles to two separate load or store instructions, so each of those two instructions from each of the two marked source lines pairwise will race (except for the two loads, which are both reads).

@@

Landslide implements both Happens-Before (henceforth referred to as {\em Pure Happens-Before} for clarity) and Limited Happens-Before.
\Cref{chap:quicksand} includes a comparison of the two approaches for the purpose of finding new preemption points for model checking.
%, we use the Limited Happens-Before relation for our analysis.
%The justification for this is that, while stand-alone data race analyses must avoid inundating the user with false alarms \cite{racerx},
%my work incorporates data race analysis in an internal feedback loop, and reports only directly observed failures to the user.
%Hence, I accept some overhead from false positives for the sake of more thorough testing.

@@

		that Landslide be only a data race detector,
